---
title: "6. Survey pages"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In all the experiments we've built so far, it has been assumed that a "trial" consists of a single stimulus to which a participant must give a response. By varying the stimulus type and the response type this produces a family of R functions that map onto different jsPsych plugins. In my experience these are the most common kinds of trials that one wishes to use, but there are also a variety of more specialised situations for which jsPsych supplies dedicated plugins. We've seen one example of this already, namely the `trial_instructions()` function. There's nothing stopping you from presenting instructions as a series of `trial_html_button_response()` trials, but it's tedious to build and it is hard (not impossible) to allow participants to browse the instruction set using forward and back buttons. So this is a situation where the specialised plugin is handy.

Another very common scenario where a specialised plugin is handy is if you want to include an entire page of survey questions. It would of course be quite easy to create a *single* survey question using `trial_html_button_response()`, and create a complete survey by using many such questions, but it is sometimes inconvenient. To make surveys a little simpler, jsPsych has plugins for five different kinds of survey page, four of which are currently supported in jaysire. The four supported types are:

- Pages where every survey question is a multiple choice question (i.e., choose exactly one option) can be created using the `trial_survey_multi_choice()` function
- Pages where every survey question is a multiple selection question (i.e., select as many options as are applicalbe) can be created using the `trial_survey_multi_select()` function
- Pages where every survey question is a Likert scale item can be created with `trial_survey_likert()`
- Pages where every survey question asks for a free text response can be created with `trial_survey_text()`

The jsPsych library does not have specialised plugins for the case where a survey page can contain questions of different types, and as a consequence neither does jaysire.


```{r example6}
library(jaysire)

page1 <- trial_survey_multi_choice(
  preamble = "Welcome! We'd like to ask some demographic questions",
  questions = list(
    question_multi(
      prompt = "Please select the option that best matches your gender",
      options = c("Male", "Female", "Nonbinary", "Other", "Prefer not to say"),
      name = "gender"
    ),
    question_multi(
      prompt = "Do you consider yourself to be LGBTIQ+?",
      options = c("Yes", "No", "Unsure", "Prefer not to say"),
    )
  )
)

page2 <- trial_survey_multi_select(
  questions = question_multi(
    prompt = "Which of the following R packages to you use?",
    options = c(
      "ggplot2", "dplyr", "purrr", "janitor", "data.table", "testthat",
      "usethis", "tibble", "magrittr", "rlang", "babynames", "janeaustenr"
    ), 
    horizontal = FALSE
  )
)


confidence_scale <- c(
  "Very unconfident", 
  "Somewhat unconfident", 
  "Somewhat confident", 
  "Very confident"
)

page3 <- trial_survey_likert(
  preamble = "How confident in you R skills?",
  questions = list(
    question_likert("Data wrangling?", confidence_scale),
    question_likert("Data visualisation?", confidence_scale),
    question_likert("Statistical modelling?", confidence_scale),
    question_likert("Designing experiments?", confidence_scale),
    question_likert("R markdown documents?", confidence_scale)
  )
)

page4 <- trial_survey_text(
  questions = question_text(
    prompt = "Anything else you would like to mention?",
    placeholder = "Type your answer here",
    rows = 8,
    columns = 60
  )
)
```



You can check out a working version of the experiment [here](../demos/example06/experiment/index.html).

```{r include=FALSE, message=FALSE, warning=FALSE}
finish <- trial_html_keyboard_response(
  stimulus = "All done! Click <a href='../../../articles/jaysire06.html'>here</a> to return to the vignette.",
  choices = no_key()
)

# write a working copy of the experiment into the docs folder
experiment(
  timeline = timeline(page1, page2, page3, page4, finish),
  path =  here::here("docs", "demos", "example06")
)
```

