---
title: "2. Randomisation, repetition and variables"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [previous article](jaysire01.html) illustrated how to build a very simple experiment using jaysire and deploy it locally. In this article I take the same experiment and elaborate it in three ways:

- Avoid code duplication using timeline variables
- Randomise the order of trials within a block
- Repetition of blocks of trials


Complete code for this demo:

```{r example2}
library(jaysire)

# ----------- instructions -----------
instructions <- trial_instructions(
  pages = c(
    "Welcome! Use the arrow buttons to browse these instructions",
    "Your task is to decide if an equation like '2 + 2 = 4' is true or false",
    "You will respond by clicking a button",
    "Press the 'Next' button to begin!"
  ),
  show_clickable_nav = TRUE,
  post_trial_gap = 1000
)


# ----------- all stimuli -----------
equations <- c(
  "13 + 23 = 36",  
  "17 - 9 = 6", 
  "125 / 5 = 25", 
  "2 - 4 = 6", 
  "12 + 39 = 43", 
  "4 * 23 = 92"
)

# ----------- trials from template -----------
trial_template <- trial_html_button_response(
  stimulus = variable("stimulus"),
  choices = c("true", "false"),
  post_trial_gap = 1000
)

trials <- timeline(trial_template) %>%
  with_variables(stimulus = equations) %>%
  with_parameters(randomize_order = TRUE, repetitions = 2)


# ----------- final trial -----------
finish <- trial_html_keyboard_response(
  stimulus = "All done! Press any key to finish",
  choices = any_key()
)

# ----------- construct timeline -----------
all_events <- timeline(
  instructions,
  trials,
  finish
)

# ----------- experiment location -----------
exp_path <- file.path(tempdir(), "equations_exp")


# ----------- build the experiment -----------
experiment(
  timeline = all_events,
  path = exp_path,
  on_finish = save_locally()
)
```


You can check out a working version of the experiment [here](../demos/example02/experiment/index.html).

```{r include=FALSE, message=FALSE, warning=FALSE}

finish2 <- trial_html_keyboard_response(
  stimulus = "All done! Click <a href='../../../articles/jaysire02.html'>here</a> to return to the vignette.",
  choices = no_key()
)

# write a working copy of the experiment into the docs folder
exp <- here::here("docs", "demos", "example02")
experiment(
  timeline = timeline(instructions, trials, finish2),
  path = exp,
)
```


